**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-28 Schloss_PacBio16S_PeerJ_2016.Rmd
Edit Conclusion section

```{r}

getRegionNonRandom <- function(region){
	file_name <- paste0("data/mothur_pool/", region, ".mock.screen.unique.good.filter.unique.error.summary")

@@

percent_one_off <- 100 * sapply(non_random, function(x){x[1]/sum(x)})

@@

## Conclusions
The various sequencing platforms that are available to microbial ecologists are
able to fill unique needs and have their own strengths and weaknesses. For
sequencing the 16S rRNA gene, the 454 platform is able to generate a moderate
number of high-quality 500-nt sequence fragments (error rates below 0.02%) {Schloss, 2011 #2466} and
the MiSeq platform is able to generate a large number of high-quality 250-nt
sequence fragments (error rates below 0.02%) {Kozich, 2013 #2719}. The promise of the PacBio
sequencing platform was the generation of high-quality near full-length sequence
fragments. As we have shown in this study, it is possible to generate near
full-length sequences; however, the error rate associated with those reads is
considerable (i.e. `r round(100*pooled_error["v19","precluster.error"],2)`%)
and requires a level of sequencing coverage that is not commonly observed in a
typical sequencing run. This results in the generation of a small number of low
quality full-length sequences. When we considered the shorter V4 region, which is
similar in length to what is sequenced by the MiSeq platform, the error rates
we observed with the PacBio platform were nearly 5-fold higher than what has
previously been reported. It appears that the promise offered by the PacBio
platform has not been realized.

The widespread adoption of the 454 and MiSeq platforms and decrease in the use
of Sanger
sequencing for the 16S rRNA gene has resulted in a decrease in the
generation of the full-length reference sequences that are needed for performing
phylogenetic analyses and designing lineage specific PCR primers and fluorescent
*in situ* hybridization (FISH) probes. It remains to be determined whether the
elevated error rates we observed for full-length sequences are prohibitive for
these applications. We can estimate the distribution of errors assuming that the
errors follow a binomial distribution along the length of the 1,500 nt
gene with the error rate that we achieved from the V1-V9 mock community data
prior to pre-clustering the sequences, which was
`r round(100*final.error["v19","unique.error"],2)`% (Figure 5). Under
these conditions one would only expect
`r round(100*dbinom(0, 1500, final.error["v19","unique.error"]), 2)`% of the
sequences to have no errors. In fact, 95% of the reads would have at least
`r qbinom(0.05, 1500, final.error["v19","unique.error"])` errors and 50% of the
reads would have at least
`r qbinom(0.50, 1500, final.error["v19","unique.error"])` errors. If the error
rate could be dropped to 0.25%, then 95% of
the reads would have at least
`r qbinom(0.05, 1500, 0.0025)` error and 50% of the
reads would have at least
`r qbinom(0.50, 1500, 0.0025)` errors. If it were possible to replicate the low
error rates we have previously observed using the 454 and Illumina MiSeq
platforms, which was 0.02%, then we would expect
`r round(100*dbinom(0, 1500, 0.0002), 1)`% of the sequences to have no errors.
In fact, 95% of the reads would have `r qbinom(0.95, 1500, 0.0002)` or fewer
errors. Although full-length sequence data is highly desired, at this point, it
does not appear that the PacBio platform can provide the data of sufficient
quality to fill the niche of generating reference sequences.

@@

The development of newer sequencing technologies continue to advance
and there is justifiable excitement to apply these technologies to sequence the
16S rRNA gene. Although it is clearly possible to generate sequencing data from
these various platforms, it is critical that we assess the platforms for their
ability to generate high quality data and the particular niche that the new
approach will fill. With this in mind, it is essential that researchers utilize
mock communities as part of their experimental design so that they can quantify
their error rates. The ability to generate large numbers of near full-length 16S
rRNA gene sequences is an exciting advance. At this point, the excitement must
be tempered by the appreciation that the error rates limit the application of
the approach.

@@

**Figure 5. The percentage of V1-V9 sequences that were predicted to have
between 0 and 20 errors as a function of the error rate of the sequences.** The
highest error rate, `r round(100*final.error["v19","unique.error"], 2)`%,
corresponds to what was observed before the pre-clustering step.
The smallest error rate (0.02%) corresponds to our previous observations using
the 454 and MiSeq sequencing platforms. The predicted number of errors was
assumed to follow a binomial distribution.

@@

### Figure 5

```{r Figure_5, fig.width=7, fig.height=4}
x <- 0:20
e1 <- final.error["v19","unique.error"]
e2 <- 0.0025
e3 <- 0.0010
e4 <- 0.0002

par(mar=c(5,6,0.5, 0.5))
plot(x, dbinom(x, 1458, e1), type="l", ylim=c(0,0.35), xlab="Number of errors",
		 ylab="Full-length 16S rRNA\ngene sequences (%)", yaxt="n", lwd=3)
points(x, dbinom(x, 1458, e2), type="l", col="red", lwd=3)
points(x, dbinom(x, 1458, e3), type="l", col="blue", lwd=3)
points(x, dbinom(x, 1458, e4), type="l", col="darkgreen", lwd=3)
axis(2, at=seq(0,0.35,0.05), label=seq(0,35,5), las=1)
legend(x=12, y=0.30, legend=paste0(format(round(100*c(e1, e2, e3, e4), 2), nsmall=2), "%"),
			 lwd=3, col=c("black", "red", "blue", "darkgreen"))
```

```{r, versions}
sessionInfo()
```

>>>>

```{r}
getRegionNonRandom <- function(region){
	file_name <- paste0("data/mothur_pool/", region, ".mock.screen.unique.good.filter.unique.error.summary")

@@

percent_one_off <- 100 * sapply(non_random, function(x){x[1]/sum(x)})
```

@@

## Conclusions
The various sequencing platforms that are available to microbial ecologists are
able to fill unique needs and have their own strengths and weaknesses. For
sequencing the 16S rRNA gene, the 454 platform is able to generate a moderate
number of high-quality 500-nt sequence fragments (error rates below 0.02%) {Schloss, 2011 #2466} and
the MiSeq platform is able to generate a large number of high-quality 250-nt
sequence fragments (error rates below 0.02%) {Kozich, 2013 #2719}. The promise
of the PacBio sequencing platform was the generation of high-quality near
full-length sequence fragments. As we have shown in this study, it is possible
to generate near full-length sequences with error rates that are slightly
higher, but comparable to the other platforms (i.e.
`r round(100*pooled_error["V1V9","precluster"],2)`%).
With the exception of the V4 region, the error rates were less than
`r 100*max(pooled_error[!(rownames(pooled_error) %in% "V4"),"precluster"])`%.
When we considered the shorter V4 region, which is similar in length to what is
sequenced by the MiSeq platform, the error rates we observed with the PacBio
platform were nearly 8-fold higher than what has previously been reported. It
was unclear why these shorter reads had such a high error rate relative to the
other regions. At this point, the primary limitation of generating full-length
sequences on the PacBio platform is the cost of generating the data and
accessibility to the sequencers.

The widespread adoption of the 454 and MiSeq platforms and decrease in the use
of Sanger sequencing for the 16S rRNA gene has resulted in a decrease in the
generation of the full-length reference sequences that are needed for performing
phylogenetic analyses and designing lineage specific PCR primers and fluorescent
*in situ* hybridization (FISH) probes. It remains to be determined whether the
error rates we observed for full-length sequences are prohibitive for
these applications. We can estimate the distribution of errors assuming that the
errors follow a binomial distribution along the length of the 1,500 nt
gene with the error rate that we achieved from the V1-V9 mock community data
prior to pre-clustering the sequences, which was
`r round(100*pooled_error["V1V9","unique"],2)`% (Figure 5). Under
these conditions one would only expect
`r round(100*dbinom(0, 1500, pooled_error["V1V9","unique"]), 2)`% of the
sequences to have no errors and 50% of the sequences would have at least
`r qbinom(0.50, 1500, pooled_error["V1V9","unique"])` errors. After applying the pre-clustering denoising step, the erro rate drops by `r round(pooled_error["V1V9","unique"]/pooled_error["V1V9","precluster"])`-fold to
`r 100*pooled_error["V1V9","precluster"]`%. With this error rate, we would
expect `r round(100*dbinom(0, 1500, pooled_error["V1V9","precluster"]), 2)`%
of the sequences to have no sequencing errors. Of course, the tradeoff is the
loss of resolution among closely related sequences.

@@

The development of newer sequencing technologies continue to advance
and there is justifiable excitement to apply these technologies to sequence the
16S rRNA gene. Although it is clearly possible to generate sequencing data from
these various platforms, it is critical that we assess the platforms for their
ability to generate high quality data and the particular niche that the new
approach will fill. With this in mind, it is essential that researchers utilize
mock communities as part of their experimental design so that they can quantify
their error rates. The ability to generate near full-length 16S rRNA gene
sequences is an exciting advance that will hopefully expand our ability to
improve the characterization of microbial communities.

@@

@@
